Proof of uniform convergence for a cell-centered AP discretization of the hyperbolic heat equation on general meshes

### 2015-03-23 Corrected_version.tex
add a remark on time rate

To illustrate the theory and have a more quantitative version of the error estimates
studied in this work, we consider the academic square $\Omega=[0,1]^2$ and discretize the hyperbolic heat equation
of a mesh made with random quads. 
A random quad mesh is made of quads where the vertices are moved
randomly around their initial position, by a factor between 10\% and 30\%.
We use the fully implicit time discretization version of the 2D scheme
detailed in this work. The solution of the linear systems is computed via
an iterative GMRES algorithm, which converges smoothly in our numerical experiments.
The reference analytical solution used in our tests is designed by separation of variables.
A solution of (\ref{eq:mod1}) is
$$
p= f+\frac{\eps^2}\sigma \partial_t f
\mbox{ and }
\mathbf u= -\frac\eps \sigma \nabla f,
$$
with $f$ solution of 
\begin{equation}\label{telegraph}
\partial_t f +\frac{\eps^2}\sigma \partial_t^2 f -\frac1\sigma \Delta f=0.
\end{equation}
 We propose to construct a solution for a subset of small $\eps$ to validate the uniform convergence. Firstly we consider that the solution is a periodic solution on the square $\left[0,2\right]\times\left[0,2\right]$. For this we use the separation of the variables. We consider the following function
$$
f(t,x,y)=\alpha(t)\cos (\pi x ))\cos( \pi y ).
$$
and we propose to find the function $\alpha(t)$ such that $f(t,x,y)=\alpha(t)\cos (\pi x ))\cos( \pi y )$ is a periodic solution of (\ref{telegraph}).
The function $\alpha$ is determined as the solution of
$$
\alpha'(t)+ \frac{\eps^2}\sigma\alpha''(t)+\frac{2\pi}\sigma \alpha=0
$$
with $\alpha'(0)=0$ and $\alpha(0)=1$.
For small $\eps$, which is the case we are interested in, the solution is computed as follows.
First determine 
$$
\lambda_1=-
\frac{
\sigma \left( \sqrt{1-\frac{\eps^2}{\sigma^2}  8 \pi^2  } +1\right)
}{2 \eps^2  }
\mbox{ and }
\lambda_2=
\frac{
\sigma \left( \sqrt{1-\frac{\eps^2}{\sigma^2}  8 \pi^2  } -1\right)
}{2 \eps^2  }.
$$
Then 
$$
\alpha(t)=\frac{\lambda_2}{\lambda_2-\lambda_1}e^{\lambda_1 t  }
-
\frac{\lambda_1}{\lambda_2-\lambda_1}e^{\lambda_2 t  }
$$
from which $p(t)$ and $\mathbf u(t)$ are easily recovered.

@@

As predicted by  the theory, the scheme is uniformly AP and the error behavior is a continuous function of $\gamma$
between the  hyperbolic and parabolic limits. However the results are much better, in the sense
the order is greater than the theoretical prediction since
the order is approximatively 1 for $\gamma=0$ (hyperbolic limit)
and 2 for $\gamma=2$ (parabolic regime).  We can find a closed result on the second order convergence for the parabolic regime in the paper \cite{TAHO} (1D linear problem).
The reason is probably that the theory is based on worst case estimates, as it is often the case for the numerical analysis
of finite volume schemes \cite{FV}.

>>>>

To illustrate the theory and have a more quantitative version of the error estimates
studied in this work, we consider the academic square $\Omega=[0,1]^2$ and discretize the hyperbolic heat equation
of a mesh made with random quads. 
A random quad mesh is made of quads where the vertices are moved
randomly around their initial position, by a factor between 10\% and 30\%.
We use the fully implicit time discretization version of the 2D scheme
detailed in this work. The solution of the linear systems is computed via
an iterative GMRES algorithm, which converges smoothly in our numerical experiments.
The reference analytical solution used in our tests is designed by separation of variables.
A solution of (\ref{eq:mod1}) is
$$
p= f+\frac{\eps^2}\sigma \partial_t f
\mbox{ and }
\mathbf u= -\frac\eps \sigma \nabla f,
$$
with $f$ solution of 
\begin{equation}\label{telegraph}
\partial_t f +\frac{\eps^2}\sigma \partial_t^2 f -\frac1\sigma \Delta f=0.
\end{equation}
 We propose to construct a solution for a subset of small $\eps$ to validate the uniform convergence. Firstly we consider that the solution is a periodic solution on the square $\left[0,2\right]\times\left[0,\frac{2}{L}\right]$. For this we use the separation of the variables. We consider the following function
$$
f(t,x,y)=\alpha(t)\cos (L \pi x ))\cos( L\pi y ).
$$
and we propose to find the function $\alpha(t)$ such that $f(t,x,y)=\alpha(t)\cos (L\pi x )\cos( L\pi y )$ is a periodic solution of (\ref{telegraph}).
The function $\alpha$ is determined as the solution of
$$
\alpha'(t)+ \frac{\eps^2}\sigma\alpha''(t)+\frac{2L^2\pi^2}\sigma \alpha(t)=0
$$
with $\alpha'(0)=0$ and $\alpha(0)=1$.
For small $\eps$, which is the case we are interested in, the solution is computed as follows.
First determine 
$$
\lambda_1=-
\frac{
\sigma \left( \sqrt{1-\frac{\eps^2}{\sigma^2}  8L^2 \pi^2  } +1\right)
}{2 \eps^2  }
\mbox{ and }
\lambda_2=
\frac{
\sigma \left( \sqrt{1-\frac{\eps^2}{\sigma^2}  8L^2 \pi^2  } -1\right)
}{2 \eps^2  }.
$$
Then 
$$
\alpha(t)=\frac{\lambda_2}{\lambda_2-\lambda_1}e^{\lambda_1 t  }
-
\frac{\lambda_1}{\lambda_2-\lambda_1}e^{\lambda_2 t  }
$$
from which $p(t)$ and $\mathbf u(t)$ are easily recovered.

@@

As predicted by  the theory, the scheme is uniformly AP and the error behavior is a continuous function of $\gamma$
between the  hyperbolic and parabolic limits. However the results are much better, in the sense
the order is greater than the theoretical prediction since
the order is approximatively 1 for $\gamma=0$ (hyperbolic limit)
and 2 for $\gamma=2$ (parabolic regime).  We can find a closed result on the second order convergence for the parabolic regime in the paper \cite{TAHO} (1D linear problem).
The reason is probably that the theory is based on worst case estimates, as it is often the case for the numerical analysis
of finite volume schemes \cite{FV}. Numerically we observe a time-linear rate for the numerical which shows that the time-exponential rate obtained using the Gronwall lemma is clearly not optimal.
